---
title: "numeric eda"
author: "Zhaoqianyu Xiong"
date: "2022-10-06"
output: html_document
---
```{r}
library(tidyverse)
```

add month variable using lubridate::floor_date(), the unit of the new variable is month.
```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

group by different variables will create different groups. For example, group by name will produce 3 groups.
```{r}
weather_df %>%
  group_by(name, month)
```

##summarize
let's group and summarize. n() to count the number.
```{r}
weather_df %>%
  group_by(name) %>%
  summarize(n_obs = n())
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(n_obs = n())
```

can we count in any other ways?
```{r}
weather_df %>%
  count(name, month)
```

don't use `table()`.  It is not a dataframe. You cannot do anything further after using table. (For sure, you can count the number with it.)
```{r}
weather_df %>%
  pull(month) %>% 
  table
```

You can have multiple summaries.
count the number of distinct values of date using codes below.
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

## a brief aside on 2*2 table.
create a new coulumn to evaluate the weather is cold or not, and omit the data about "WHA". Finally, group by name and cold, and count the number of each group.
```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

change the form of the data.
```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(
    n_obs = n()
  ) %>%
  pivot_wider(
    names_from = cold,
    values_from = n_obs
  )
```


```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ "" ##if the above lines does not work out (nothing else's true), then the value of the row will be" ".
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold) ##create a 2*2 table quickly by janitor::tabyl
```

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_path()
```

create table using knitr:kable(). digit = "" is the number of decimal.
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

## Grouped mutate
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax))
```

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>%
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

lagged observations (指定行的前一行数值)
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    yesterday_tmax = lag(tmax)) %>%
  view()
```

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```

One other window function
```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(tmax) < 2)
```


