---
title: "viz_part_1"
author: "xiong zqy"
output: html_document
date: "2022-09-29"
---
#import data
```{r}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(     ##download NOAA weather data
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Let's make a scatter plot.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```
only make a plot.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
Let's make a same scatterplot, but different.
```{r}
weather_df %>%
  drop_na %>% 
  filter(name == "CentralPark_NY") %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Let' keep making a same scatterplot, but different. (the plot is saved.)
```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```

## let's fancy this up a bit.
```{r}
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth()
```
按名字区分color 仅在点图中， 因为只在geom_point()
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes()) +
  geom_smooth()
```
write 'color' when making plot (the first step), it will be followed in every graphs.
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)
```
'alpha' make the graph transparent, 

make separate panels. (by facet_grid)
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~name)
```
纵向

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~.)
```
横向

'tmax' vs 'tmin' is boring, let's spice it up some.
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) +
  theme(axis.text.x = element_text(angle = 90)) ##rotate the x axis 90'
```

## some quick stuff
```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```
when the point is too dense, use hex

histograms, barplots, boxplots, violins, ...
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```
"dodge" separate the bars

MORE OPTIONS!
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

boxplot
```{r}
weather_df %>%
ggplot(aes(x = name, y = tmax)) + geom_boxplot()
```

violin plots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

density ridges
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges()
```

distribution (?scale)

##saving and emdedding plots.
First, let's save a plot.
```{r}
weather_scatterplot = 
weather_df %>%
ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) +
  theme(axis.text.x = element_text(angle = 90))

weather_scatterplot

ggsave("./results/weather scatterplot.pdf", weather_scatterplot,
       width = 8, height = 5)
```

```{r}
weather_scatterplot
```

